---
title: "Oakland Parcel Analysis"
author: "Robert Spragg"
date: "7/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(leaflet)
if (!require(leaflet.mapboxgl)) {
  devtools::install_github("rstudio/leaflet.mapboxgl")
  library(leaflet.mapboxgl)
}
library(sf)
```


```{r}
path <-
  "/Users/robert/ab-1279/Alameda/Parcels-shp/"
```

```{r}
alameda_county_parcel_data <-
  read_sf(path)
```

Constants
```{r}
QUARTER_ACRE <-
  10890 # SQUARE FEET
HALF_ACRE <-
  21780 # SQUARE FEET
NONZERO_PARCEL_SIZE_COUNT_UNIQUE_APN_SORT <-
  97474

access_token <-
  "pk.eyJ1Ijoicm9iZXJ0c3ByYWdnIiwiYSI6ImNqd2NtNjh0YzBiNzMzenBsNDl3cXQ1em8ifQ.1d2QhIjr8R8ZUlzjwcYx9w"

style_oakland_map <-
  "mapbox://styles/robertspragg/ckce9o0fi09vb1ipdrpo41iq8"
```

Load the Alameda County Parcel shapefile, select metadata of interest, and transform coordinates.
```{r}
parcels_geom_only <-
  alameda_county_parcel_data %>% 
  select(geometry, APN_SORT, OBJECTID, LotSize, SitusAddre) %>% 
  st_transform(crs = 4326)
```


To find the Oakland parcels of interest, first load the Alameda County shapefile, remove the geometry to speed things up, and filter for all Oakland parcels.
```{r}
parcels_no_geometry <-
  alameda_county_parcel_data %>% 
  st_drop_geometry()

oakland_parcels <-
  parcels_no_geometry %>% 
  filter(SitusCity == "OAKLAND")
```



Next, let's find the number of Oakland parcels larger than a quarter acre (10890 square feet)
```{r}
oakland_parcels %>% 
  group_by(APN_SORT) %>% 
  slice(1) %>%  # 98,294
  filter(LotSize > 0) %>% # 97,474
  filter(LotSize > QUARTER_ACRE) %>% 
  nrow()

```

Let's repeat this for parcels greater than half an acre.
```{r}
oakland_parcels %>% 
  group_by(APN_SORT) %>% 
  slice(1) %>%  # 98,294
  filter(LotSize > 0) %>% # 97,474
  filter(LotSize > HALF_ACRE) %>% 
  nrow()
```

There are 10,723 parcels greater than a quarter acre, and 4,245 greater than a quarter acre.
Now, let's save the lists of these parcels.
```{r}
parcel_list_half_acre <-
  oakland_parcels %>% 
  group_by(APN_SORT) %>% 
  slice(1) %>%  # 98,294
  filter(LotSize > 0) %>% # 97,474
  filter(LotSize > HALF_ACRE) %>% 
  select(APN_SORT, LotSize) %>% 
  select(APN_SORT) %>% pull

```

```{r}
parcel_list_quarter_acre <-
  oakland_parcels %>% 
  group_by(APN_SORT) %>% 
  slice(1) %>%  # 98,294
  filter(LotSize > 0) %>% # 97,474
  filter(LotSize > QUARTER_ACRE) %>% 
  select(APN_SORT, LotSize) %>% 
  select(APN_SORT) %>% pull

```

Finally, using our original shapefile data, let's plot the parcels whose APNs are in the list of parcels greater than half an acre in size.
```{r}
leaflet(parcels_geom_only %>% filter(APN_SORT %in% parcel_list_half_acre)) %>% 
    addMapboxGL(
    accessToken = access_token,
    style = style_oakland_map
  ) %>% 
  addPolygons(color = "#444444", weight = 1, smoothFactor = 0.5,
    opacity = 1.0, fillOpacity = 0.5,
    fillColor = ~colorQuantile("YlOrRd", LotSize)(LotSize),
    label = ~SitusAddre,
    highlightOptions = highlightOptions(color = "white", weight = 2,
      bringToFront = TRUE))
```
